Theorems and Definitions

### 2016-04-07 number_theory/hw/hw9.tex
finished pres for now

\item Suppose that $n$ is a positive integer and $k$ is any integer.
\begin{enumerate}
\item Prove that $gcd(n,k)=1$ if and only if $gcd(n,n-k)=1$.
\begin{quote}
$\Leftarrow)$ Suppose that $gcd(n,k)>1$. Then $gcd(n,k)=q$, so $q\mid n$ and
$q\mid k$, $n=qa$ and $k=qb$ for some $a,b\in\ZZ$. Then $n-k=pa-pb=p(a-b)$, so $p\mid n-k$, which
implies that $gcd(n,n-k)>1$. Therefore, by contrapositive, $gcd(n,k)=1$ implies $gcd(n,n-k)=1$.
\noindent\\\\
$\Rightarrow)$ Suppose $gcd(n,n-k)>1$. Then there exists $p\in\ZZ$ such that $p\mid n$ and $p\mid k$.
Then $n=pa$ and $k=pb$ for some $a,b\in\ZZ$. We can express $n-k=pa-pb=p(a-b)$, so $p\mid(n-k)$ and
$p\mid n$, and $gcd(n,n-k)>1$.
\end{quote}
\item Prove that $\phi(n)$ is an even integer for all $n\geq3$.
\begin{quote}
By definition, the totient function counts the number of units in $\ZZ_{n}$. Suppose that
$n\geq3$. If we take any element $k$ from $\ZZ_{n}$, we know by $(a)$ that if $k$ is relatively
prime to $n$, then $n-k$ is relatively prime to $n$. Therefore, if $k$ is a unit in $\ZZ_{n}$, then
$n-k$ is also a unit in $\ZZ_{n}$. Then for every unit $k$ in $\ZZ_{n}$, we can find another unit
$n-k$ also in $\ZZ_{n}$, which means there are an even number of units in $\ZZ_{n}$, so
$\phi(n)\in\{ 2j\ |\ j\in\ZZ \}$.
\end{quote}
\end{enumerate}

>>>>

\item Suppose that $n$ is a positive integer and $k$ is any integer.
\begin{enumerate}
\item Prove that $gcd(n,k)=1$ if and only if $gcd(n,n-k)=1$.
\begin{quote}
$\Leftarrow)$ Suppose that $gcd(n,k)>1$. Then $gcd(n,k)=q$, so $q\mid n$ and
$q\mid k$, $n=qa$ and $k=qb$ for some $a,b\in\ZZ$. Then $n-k=pa-pb=p(a-b)$, so $p\mid n-k$, which
implies that $gcd(n,n-k)>1$. Therefore, by contrapositive, $gcd(n,k)=1$ implies $gcd(n,n-k)=1$.
\noindent\\\\
$\Rightarrow)$ Suppose $gcd(n,n-k)>1$. Then there exists $p\in\ZZ$ such that $p\mid n$ and $p\mid k$.
Then $n=pa$ and $k=pb$ for some $a,b\in\ZZ$. We can express $n-k=pa-pb=p(a-b)$, so $p\mid(n-k)$ and
$p\mid n$, and $gcd(n,n-k)>1$.
\end{quote}
\item Prove that $\phi(n)$ is an even integer for all $n\geq3$.
\begin{quote}
By definition, the totient function counts the number of units in $\ZZ_{n}$. Suppose that
$n\geq3$. If we take any element $k$ from $\ZZ_{n}$, we know by $(a)$ that if $k$ is relatively
prime to $n$, then $n-k$ is relatively prime to $n$. Therefore, if $k$ is a unit in $\ZZ_{n}$, then
$n-k$ is also a unit in $\ZZ_{n}$. If $n$ is odd, then we know that there will not be any situation
where $n-k=k$, since that would imply that $n=2k$, or $n$ is even which is a contradiction. Then for
every unit $k$ in $\ZZ_{n}$, we can find another unit $n-k$ also in $\ZZ_{n}$, which means there are
an even number of units in $\ZZ_{n}$, so $\phi(n)\in\{ 2j\ |\ j\in\ZZ \}$ when $n$ is odd. If $n$
is even, however, there may be a unit $k$ in $\ZZ_{n}$ such that $k=n-k$. If this is the case,
then this unit will be the same as $n/2$. Since it is a unit, then $gcd(n/2,n)=1$. We saw that
$n$ was even, so we can say $n=2l,\ l\in\ZZ$. Then $gcd(2l/2,2l)=gcd(l,2l)\neq1$, so there cannot
be a unit $k$ in an even $\ZZ_{n}$ such that $k=n-k$.
\end{quote}
\end{enumerate}
